---
title: 'Dealing with database'
author: "Ebude Yolande"
output:
  pdf_document: default
  html_notebook: default
---

# Introduction
Nowadays, most companies use database to store information. There are different kind of database but in this workshop we will handle Structured Query Language (SQL) type: SQLite. To start download these datasets

- \href{https://drive.google.com/file/d/151p846LIvDjQD9WG3fRu2w5L2rQFZSK2/view?usp=sharing}{SqLite file}

Install Rstudio \href{https://rstudio.com/products/rstudio/download/}{\textit{(download)}} or update version
```{r}
#install.packages("installr")
#library(installr)
#updateR()
```

# Content
- Import different database types
- Modify database 

# 1. Import different database types

## SQLite
Install sqlite in computer \href{https://www.sqlite.org/download.html}{download file}.
```{r}
library(RSQLite)
conn <- dbConnect(SQLite(),'GeneratorRefilling.sqlite')
dtables<-dbListTables(conn) # list tables in db
query <- 'SELECT * From RefillingGen2018'
dGenR<- dbGetQuery(conn,query)
#tail(dGenR,3)
```

## Dat files
```{r}
tab<-read.delim2('3055.dat')
temp<-colnames(tab)
colnames(tab)<-c('num','val','Firstname','Lastname','Email','val1','unit','Date','Datetime','Verify')
head(tab,3)
```


# 2 Modify database
Create new table
```{r}
#query<- 'CREATE TABLE Workshop (id integer PRIMARYKEY, name char(25), time Date)'
#dbSendQuery(conn, query)
```
Create new table of an existing dataframe
```{r}
dbCreateTable(conn,'Tab3055',tab)
dtables<-dbListTables(conn)
dtables
```

Add rows in new table
```{r}
query<-"INSERT INTO Workshop(id, name, time) VALUES ('1','Penina','12/01/2019')"
dbSendQuery(conn,query)
```
Delete rows in table
```{r}
query<-"DELETE FROM Tab3055 WHERE Firstname=='Mary'"
dbExecute(conn,query)
```
Change a value in the table
```{r}
query<-"UPDATE Workshop SET name='Angela' WHERE id='1'" 
dbExecute(conn,query)  
```
Check values of table
```{r}
query<-"SELECT * FROM Workshop"
dbtab<-dbGetQuery(conn, query)
#dbtab
```

close the database
```{r}
dbDisconnect(conn)
```


# Challenge

1. Import "GeneratorRefilling.sqlite" and "3057.dat" and change the header of "3057.dat" to ['num','name','surname','date']

2. Create a table in the database for "3057.dat" called Tab3057

3. Delete the "Tab3055" table

4. Add a row to Tab3057 table

5. Delete the 3rd row of Tab3057

6. Add column 'Time' to Tab3057
